Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-10 doc/manuscript/translocation_pnas.tex
Reduce viability methods and address Max's comments

\title{Reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease}

@@

There are four conditions that will often determine the feasibility of
using individuals from populations that are recovering following
disease-induced declines to reestablish extirpated populations. First,
populations that survive following disease emergence evolve increased
resistance/tolerance. Second, increased resistance/tolerance contributes
to higher individual fitness in the presence of disease and allows
positive population growth. Third, the fitness benefits gained through
selection for increased resistance are maintained beyond the natal
environment and across generations, allowing for the establishment of
reintroduced populations. And fourth, reintroduced populations show
long-term growth and persistence. To date, few of these conditions are
documented for any single species impacted by disease, and as a result
the reintroduction of resistant hosts as a means to recover species of
conservation concern remains largely an untested idea
\citep{scheele2021, mendelson2019}.

The mountain yellow-legged (MYL) frog, comprised of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} \citep{vredenburg2007}, is
emblematic of the unprecedented global declines of amphibians caused by
the recently-emerged amphibian chytrid fungus
\citep[\emph{Batrachochytrium dendrobatidis,}][]{scheele2019}. Once the
most common amphibian in the high elevation portion of the California's
Sierra Nevada mountains \citep[USA,][]{grinnell1924}, this frog has
disappeared from more than 90\% of its historical range during the past
century \citep{vredenburg2007}. Due to the severity of its decline and
the increasing probability of extinction, both species are now listed as
``endangered'' under the U.S. Endangered Species Act. In the Sierra
Nevada, this decline was initiated by the introduction of non-native
trout into the extensive historically-fishless region
\citep{bradford1989, knapp2000}, and the arrival of \emph{B.
dendrobatidis} (Bd) in the 1960s and its subsequent spread
\citep{vredenburg2019} caused additional widespread population
extirpations \citep{vredenburg2010, rachowicz2006}. These Bd-caused
declines are fundamentally different from the fish-caused declines
because removal of fish populations is feasible and results in the rapid
recovery of frogs \citep{knapp2007, vredenburg2004}. In contrast, Bd
appears to persist in habitats even in the absence of amphibian hosts
\citep{walker2007}, and therefore represents a long-term alteration of
invaded ecosystems that amphibians will need to overcome to reestablish
populations.

Despite the catastrophic impact of Bd on MYL frogs in which most
Bd-naive populations are driven to extirpation following Bd arrival
\citep{vredenburg2010}, some populations have persisted after epizootics
\citep{briggs2010} and are now recovering \citep{knapp2016}. Frogs in
these recovering populations show reduced susceptibility to Bd
infection, consistent with selection for more resistant/tolerant
genotypes \citep{knapp2016}. Furthermore, exploratory efforts to
reintroduce frogs collected from recovering populations to sites from
which frogs were previously extirpated indicate that population
establishment is possible despite ongoing Bd infection
\citep{joseph2018}.

@@

We conducted a principal component analysis (PCA) of the genomic data to
describe the association between sampled populations, and then used two
complementary approaches to identify regions of the genome under
selection. First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
specific single nucleotide polymorphisms (SNPs). Second, we used a
splined window analysis to identify larger genomic regions showing
differences between population types in \emph{F\textsubscript{ST}} and
nucleotide diversity (\(\pi_{diff}\) = \(\pi_{naive}\) -
\(\pi_{recovering}\)).

@@

Results from the SNP and splined window analyses show that recovering
populations have signatures of strong selection on multiple regions of
the genome. The SNP analysis identified 11 outlier SNPs from 7 distinct
genes across 4 contigs (Figure S1B,C, SI
Dataset 1). One of the 7 identified genes did not have an associated
annotation. Frequency differences between the naive and recovering
populations ranged from 41\% to 86\%. Most of these SNPs showed
frequency differences in only a subset of the sampled populations
(Figure~\ref{fig-allelefrequencies} A, B), but the outlier SNP in the
RIN3 gene showed consistent differences in frequencies across all
populations (Figure~\ref{fig-allelefrequencies} C). This is suggestive
of strong selection at this locus associated with exposure of frogs to
Bd.

The splined window analysis identified 58 outlier regions for
\emph{F\textsubscript{ST}} and 33 outlier regions for \(\pi_{diff}\)
(Figure~\ref{fig-spline-manhattan} A, B). Of these, 9 regions were
outliers for both metrics and 2 of these shared regions also contained
outlier SNPs. A total of 33 annotated genes were found in the 9 shared
regions (SI Dataset 2). The largest \(\pi_{diff}\), indicative of
directional selection, occurred in a 163kb region on Contig19, 12.9Mb
upstream of the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan} C).
This region contains approximately 500 SNPs and one annotated gene
called ``interferon-induced very large GTPase 1-like'' (GVINP1).
Additionally, a shared outlier region on Contig1 contained two
complement factor genes (C6 and C7). Interestingly, this region had a
low \(\pi_{diff}\) value, consistent with balancing selection. Finally,
one shared outlier region on Contig8 contained one outlier SNP (TCF19)
and was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan} D, Figure S2).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended MHC Class I region (FLOT1, TUBB, MDC1,
CCHCR1, TCF19) and three occur in the extended MHC Class III region
(HSP70, LSM2, VARS) \citep{ohta2006}. Therefore, this region under
selection is part of the extended MHC Class I and III complex and shows
synteny with other amphibian genomes.

@@

In this section, we describe the outcome of \emph{R. sierrae}
reintroductions conducted in Yosemite National Park as part of a 15-year
effort to reestablish populations of this endangered species
(2006--2020). Each of the 24 reintroductions involved collection of
adult frogs from 1 of 3 recovering, Bd-positive ``donor'' populations
(Figure S3) and translocating them to 1 of 12 nearby
recipient sites \citep{seddon2014}. The donor populations included 2 of
the 5 recovering populations in the frog evolution study described above
(Figure~\ref{fig-allelefrequencies} A: population 1 and 4). Following
translocation, adult survival and recruitment of new adults were
estimated from CMR surveys and counts of tadpoles and juveniles were
obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1 year following translocation
indicate that survival was highly variable between recipient sites, but
relatively constant within recipient sites (for the subset of sites that
received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability over time to choose recipient sites
with higher habitat quality for \emph{R. sierrae}. Finally, Bd loads
changed little between immediately prior to translocation versus during
the one year after translocation (Figure S5), and loads were nearly always well below the level indicative of
severe chytridiomycosis and associated frog mortality (5.8 ITS copies on
a log\textsubscript{10} scale) \citep{vredenburg2010, joseph2018}.

@@

Results from frog translocations indicated that adult survival was often
relatively high and that most populations showed evidence of successful
reproduction and recruitment (described above). Although suggestive of
population establishment, a decade or more of surveys may be necessary
to confirm that populations are in fact self-sustaining
\citep{joseph2018}. To extend our inferences of population establishment
beyond those possible from the 2-16 years of site-specific CMR data, we
developed a population viability model. Specifically, to test whether
the observed yearly adult survival probabilities in translocated
populations were sufficient to promote long-term viability, we built a
stage-structured matrix model that captured known frog demography and
included demographic and environmental stochasticity. We parameterized
the model using CMR data from translocated populations and known life
history values in this system (Table S2). Given
observed yearly adult survival probabilities (from analysis of CMR data)
and a baseline yearly recruitment probability \(\omega\) of 0.3 (i.e.,
the probability that, following survival, juveniles successfully recruit
to the adult class), our stage-structured model predicted that six of
twelve translocated populations should experience a long-run growth rate
\(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability} A; median predicted \(\lambda\) ranges from
1.15-1.36 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. When we reduced the
probability of successful recruitment \(\omega\) to 0.1 (a highly
conservative estimate), long-run growth rate of five of these six
populations was still greater than 1 (Figure~\ref{fig-viability} A). In
contrast, for the other six translocated populations, where yearly adult
survival was substantially less than 0.5, our model predicted that it
was unlikely that long-run growth rate in these populations was greater
than 1, even when the probability of successful recruitment to the adult
life stage was near 0.4 (Figure~\ref{fig-viability} A). A sensitivity
analysis of long-run growth rate \(\lambda\) to yearly adult survival
probability of naturally recruited individuals \(\sigma_{A_R}\),
recruitment probability \(\omega\), and female fecundity \emph{F} showed
that \(\lambda\) was most sensitive to small changes in \(\sigma_{A_R}\)
(Figure~\ref{fig-viability-supp} A). Note that we assumed
\(\sigma_{A_R}\) was equivalent to the survival probability of
translocated adults \(\sigma_{A_T}\), except for three populations where
we had \textgreater{} 5 years of CMR data and recovering populations
that together allowed us to estimate \(\sigma_{A_R}\) and
\(\sigma_{A_T}\) separately (Figure S7).

We then incorporated demographic stochasticity (i.e., probabilistic
births and deaths of individuals in a population) and environmental
variability in the probability of juvenile recruitment to the adult
class \(\omega\) (the transition that we expect to be the most subject
to environmental stochasticity in the presence of Bd) into the model
described above and used it to predict long-term population viability.
Our model predicted that, following a single introduction of 40 adult
individuals into a population and using the baseline value for
\(\omega\) of 0.3, the six populations with the highest adult survival
probabilities had 50-year extinction probabilities of 0 to 0.32;
Figure~\ref{fig-viability} B). This indicates strong potential for
long-term persistence in the presence of Bd and environmental
variability in recruitment (Figure~\ref{fig-viability} B). When we
decreased \(\omega\) from 0.3 to a conservative estimate of 0.15, 50
year extinction probability remained less than 0.5 for three populations
and was 0.79 to 0.94 for three additional populations
(Figure~\ref{fig-viability} B). In contrast, for the six populations
where \(\sigma_{A_R} < 0.5\), extinction probability over 50 years was
always predicted to be 1. To test the validity of our model predictions,
we demonstrated that our stochastic model could describe the general
recovery trajectory of our translocated population with the longest
survey history (Figure~\ref{fig-viability-supp} B; population 70550,
surveyed for 16 years).

@@

Disease-induced population declines are impacting global biodiversity
\citep{daszak2000}, and strategies to recover affected species are
lacking. Recovery in this context is likely to be particularly
challenging because typically recovery will need to occur in the
presence of ongoing disease. One potential strategy, and the one we
employ in the current study, utilizes populations of resistant/tolerant
individuals as a source of animals with which to reestablish extirpated
populations in the presence of disease. Although previously proposed
\citep{mendelson2019}, this strategy has rarely been successfully
implemented (see \citep{joseph2018} for a small-scale example). In
particular, although numerous studies have documented signatures of
selection in hosts following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (i) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease (Bd) and that some
of the genomic regions under selection are associated with genes related
to disease resistance/tolerance, and (ii) reintroduction of individuals
from these resistant/tolerant populations allows the successful
reestablishment of viable frog populations despite ongoing Bd infection.
Although these results are generally consistent with expectations under
evolutionary rescue \citep{carlson2014}, obtaining strong evidence in
wild animal populations for each of the associated causal links (i.e.,
selection for adaptive alleles \(\rightarrow\) increase in host fitness
in the presence of the pathogen \(\rightarrow\) positive population
growth and recovery) is difficult and perhaps impossible. Nonetheless,
in the following narrative, we use evolutionary rescue as an
informative, albeit tentative framework.

@@

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
substantial differences in allele frequencies and larger genomic regions
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the major
histocompatibility complex (MHC) plays an important role in immunity. In
our study, we identified a region that shows evidence of selection in
recovering populations and contains eight genes associated with either
the MHC Class I or Class III regions. These results corroborate numerous
previous studies linking MHC genes to amphibian resistance/tolerance
against Bd \citep[e.g.,][]{savage2011, bataille2015}. Similarly, the
region with the strongest indication of directional selection contains
the gene, GVINP1. Several previous studies have found this gene to be
differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018, ellison2014}. This gene was also strongly
linked to disease in salmon, explaining a remarkable 20\% of the
resistance phenotype \citep{robledo2020, robledo2018}. Finally, the RIN3
gene showed a consistent pattern of allele frequency differences across
all nine of the frog populations sampled in this study, indicating
consistent selection in populations distributed across a wide geographic
area. In the SI, we provide additional details on these and other genes
identified in our study as under selection . Collectively, these results
suggest that the documented natural recovery of MYL frogs following
Bd-caused declines \citep{knapp2016} is consistent with evolutionary
rescue.

@@

The recent emergence of Bd worldwide has contributed to the decline of
more than 500 amphibian species, nearly 100 of which now are extinct in
the wild \citep{scheele2019}. This unprecedented impact on global
amphibian biodiversity is compounded by the lack of any effective and
broadly applicable strategies to reverse these impacts
\citep{garner2016, knapp2022}. Importantly, in addition to the natural
recovery documented for MYL frogs \citep{knapp2016}, other species are
also showing evidence of post-epizootic recovery in the presence of Bd
\citep{scheele2017, voyles2018} and suggest the possibility of using
animals from these recovering populations to reestablish extirpated
populations. As with MYL frogs, the feasibility and long-term success of
such efforts will depend on the availability of robust donor populations
containing individuals that have acquired via evolution the adaptive
alleles necessary to allow frog survival and population growth in the
presence of Bd. Despite the hopeful example of successful
reestablishment of MYL frogs via evolutionary rescue and
reintroductions, the challenge of recovering hundreds of Bd-impacted
amphibians worldwide is a daunting prospect. Although we now have a
proven strategy to reestablish extirpated MYL frog populations, recovery
across their large historical range will require substantial resources
over many decades. The results of this study provide a hopeful starting
point for that endeavor, and other future efforts worldwide.

@@

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site \(\pi\) separately for individuals from the naive
and recovering populations using VCFtools, then calculated
\(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome. We then used the WINGEN R package
\citep{bishop2023} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the fitted spline, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015} . This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for F\textsubscript{ST} or above/below the mean for \(\pi_{diff}\).
Outlier regions were then identified as those that were outliers in both
analyses, meaning that they showed 1) high differentiation between naive
and recovering populations, and 2) differential patterns of nucleotide
diversity in the same region. Finally, we extracted gene transcripts
mapped within each region and retrieved annotation for that region using
blast.

@@

To assess the magnitude of changes in Bd load on frogs following
translocation, we compared Bd loads measured before versus after
translocation. Before-translocation loads were quantified using skin
swabs collected from all to-be-translocated frogs at the donor site on
the day before or the day of the translocation. After-translocation Bd
loads were based on all swabs collected from translocated frogs at the
recipient site in the year of and the year following translocation.
Individual frogs and their associated Bd loads were included in the
dataset only if frogs were captured at the recipient site at least once
during the year 0-1 period. We used an analysis workflow similar to that
described above. Briefly, the response variable was before-translocation
load and the predictor variable was after-translocation load. We
compared models with and without group-level effects, and group-level
effects included pit\_tag\_id, pit\_tag\_id nested within site\_id, or
pit\_tag\_id nested within translocation\_id. All models used a negative
binomial family. Priors, chains, iterations, and checks were as
described above. Code to create the dataset and fit models is available
from the following GitHub repository:
\url{https://github.com/SNARL1/translocation}.

@@

\subsubsection*{Incorporating yearly variability in vital rates}

>>>>

\title{Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease}

@@

There are four conditions that will often determine the feasibility of
using individuals from populations that are recovering following
disease-induced declines to reestablish extirpated populations. First,
populations that survive following disease emergence evolve increased
resistance/tolerance. Second, increased resistance/tolerance contributes
to higher individual fitness in the presence of disease and allows
positive population growth. Third, the fitness benefits gained through
selection are maintained beyond the natal environment and across
generations, allowing for the establishment of reintroduced populations.
And fourth, reintroduced populations show long-term growth and
persistence. To date, few of these conditions are documented, and
reintroducing hosts from recovering populations is largely untested
\citep{scheele2021, mendelson2019}.

The mountain yellow-legged (MYL) frog, composed of the sister species
\emph{Rana muscosa} and \emph{Rana sierrae} \citep{vredenburg2007}, is
emblematic of the unprecedented global declines of amphibians caused by
the recently-emerged amphibian chytrid fungus
\citep[\emph{Batrachochytrium dendrobatidis,}][]{scheele2019}. Once the
most common amphibian in the high elevation portion of the California's
Sierra Nevada mountains \citep[USA,][]{grinnell1924}, this frog has
disappeared from more than 90\% of its historical range during the past
century \citep{vredenburg2007}. Due to the severity of its decline and
the increasing probability of extinction, both species are now listed as
``endangered'' under the U.S. Endangered Species Act. In the Sierra
Nevada, this decline was initiated by the introduction of non-native
trout into the extensive historically-fishless region
\citep{bradford1989, knapp2000}, and the arrival of \emph{B.
dendrobatidis} (Bd) in the 1960s and its subsequent spread
\citep{vredenburg2019} caused additional widespread population
extirpations \citep{vredenburg2010, rachowicz2006}. These Bd-caused
declines are fundamentally different from the fish-caused declines
because removal of fish populations is feasible and results in the rapid
recovery of frogs \citep{knapp2007, vredenburg2004}. In contrast, Bd
appears to persist in habitats even in the absence of amphibian hosts
\citep{walker2007}, and therefore represents a long-term alteration of
invaded ecosystems that amphibians will need to overcome to reestablish
populations.

Despite the catastrophic impact of Bd on MYL frogs in which most
Bd-naive populations are extirpated following Bd arrival
\citep{vredenburg2010}, some populations have persisted after epizootics
\citep{briggs2010} and are now recovering \citep{knapp2016}. Frogs in
these recovering populations show reduced susceptibility to Bd
infection, consistent with selection for more resistant/tolerant
genotypes \citep{knapp2016}. Furthermore, exploratory efforts to
reintroduce frogs collected from recovering populations to sites from
which frogs were previously extirpated indicate that population
establishment is possible despite ongoing Bd infection
\citep{joseph2018}.

@@

We conducted a principal component analysis (PCA) of the genomic data to
describe the association between sampled populations, and then used two
complementary approaches to identify regions of the genome under
selection. First, we used a multivariate linear mixed model to evaluate
associations between population type (i.e., naive versus recovering) and
specific variants, including single nucleotide polymorphisms (SNPs) and
insertions/deletions (INDELS). Second, we used a splined window analysis
to identify larger genomic regions showing differences between
population types in \emph{F\textsubscript{ST}} and nucleotide diversity
(\(\pi_{diff}\) = \(\pi_{naive}\) - \(\pi_{recovering}\)).

@@

Results from the SNP and splined window analyses show that recovering
populations have signatures of strong selection on multiple regions of
the genome. The SNP analysis identified 11 outlier SNPs from 7 distinct
genes across 4 contigs (Figure S1B,C, SI
Dataset 1). One of the 7 identified genes did not have an associated
annotation. Frequency differences between the naive and recovering
populations ranged from 41\% to 86\%. Most of these SNPs showed
frequency differences in only a subset of the sampled populations
(Figure~\ref{fig-allelefrequencies} A, B), but the outlier SNP in the
RIN3 gene showed consistent differences in frequencies across all
populations (Figure~\ref{fig-allelefrequencies} C). This is suggestive
of strong - and parallel - selection at this locus associated with
exposure of frogs to Bd across multiple populations.

The splined window analysis identified 58 outlier regions for
\emph{F\textsubscript{ST}} and 33 outlier regions for \(\pi_{diff}\)
(Figure~\ref{fig-spline-manhattan} A, B). Of these, 9 regions were
outliers for both metrics and 2 of these shared regions also contained
outlier SNPs. A total of 33 annotated genes were found in the 9 shared
regions (SI Dataset 2). The largest \(\pi_{diff}\), indicative of
directional selection, occurred in a 163kb region on Contig19, 12.9Mb
upstream of the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan} C).
This region contains approximately 500 SNPs and one annotated gene
called ``interferon-induced very large GTPase 1-like'' (GVINP1).
Additionally, a shared outlier region on Contig1 contained two
complement factor genes (C6 and C7). Interestingly, this region had a
low \(\pi_{diff}\) value, consistent with balancing selection. Finally,
one shared outlier region on Contig8 contained one outlier SNP (TCF19)
and was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan} D, Figure S2).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether \emph{R. sierrae} from recovering populations could
be used to reestablish extirpated populations, we conducted 24
reintroductions in Yosemite National Park (2006--2020). Each of the
reintroductions involved collection of adult frogs from 1 of 3
recovering, Bd-positive ``donor'' populations (Figure S3) and translocating them to 1 of 12 nearby recipient sites
\citep{seddon2014}. The donor populations included 2 of the 5 recovering
populations in the frog evolution study described above
(Figure~\ref{fig-allelefrequencies} A: population 1 and 4). Following
translocation, adult survival and recruitment of new adults were
estimated from CMR surveys and counts of tadpoles and juveniles were
obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1 year following translocation
indicate that survival was highly variable between recipient sites, but
relatively constant within recipient sites (for the subset of sites that
received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability over time to choose recipient sites
with higher habitat quality for \emph{R. sierrae}. Finally, Bd loads
were fairly consistent before versus after translocation and loads were
nearly always well below the level indicative of severe chytridiomycosis
and associated frog mortality (Figure S5,
see SI text for additional details) \citep{joseph2018, vredenburg2010}.

@@

Results from frog translocations indicated that adult survival was often
relatively high and that most populations showed evidence of successful
reproduction and recruitment (described above). Although suggestive of
population establishment, a decade or more of surveys may be necessary
to confirm that populations are in fact self-sustaining
\citep{joseph2018}. To extend our inferences of population establishment
beyond those possible from the 2-16 years of site-specific CMR data, we
developed a population viability model. Specifically, to test whether
the observed yearly adult survival probabilities in translocated
populations were sufficient to promote long-term viability, we built a
stage-structured matrix model that captured known frog demography and
included demographic and environmental stochasticity. We parameterized
the model using CMR data from translocated populations and known life
history values in this system (Table S2).

Given observed yearly adult survival probabilities (from analysis of CMR
data) and a baseline yearly recruitment probability \(\omega\) of 0.3
(i.e., the probability that, following survival, juveniles successfully
recruit to the adult class), our stage-structured model predicted that
six of twelve translocated populations should experience a long-run
growth rate \(\lambda\) greater than 1 in the presence of Bd
(Figure~\ref{fig-viability} A; median predicted \(\lambda\) ranges from
1.15-1.36 for these six populations). These six populations all had
observed yearly adult survival greater than 0.5. When we reduced the
probability of successful recruitment \(\omega\) to 0.1 (a highly
conservative estimate), long-run growth rate of five of these six
populations was still greater than 1 (Figure~\ref{fig-viability} A).

Even when incorporating demographic and environmental stochasticity in
the probability of juvenile recruitment to the adult class \(\omega\)
(the transition that we expect to be the most subject to environmental
stochasticity in the presence of Bd), populations with high adult
survival are likely to persist over a 50 year time horizon. Our model
predicted that, following a single introduction of 40 adult individuals
into a population and using the baseline value for \(\omega\) of 0.3,
the six populations with the highest adult survival probabilities had
50-year extinction probabilities of 0 to 0.32;
Figure~\ref{fig-viability} B). This indicates strong potential for
long-term persistence in the presence of Bd and environmental
variability in recruitment (Figure~\ref{fig-viability} B). When we
decreased \(\omega\) from 0.3 to a conservative estimate of 0.15, 50
year extinction probability remained less than 0.5 for three populations
and was 0.79 to 0.94 for three additional populations
(Figure~\ref{fig-viability} B). In contrast, for the six populations
where \(\sigma_{A_R} < 0.5\), extinction probability over 50 years was
always predicted to be 1. To test the validity of our model predictions,
we demonstrated that our stochastic model could describe the general
recovery trajectory of our translocated population with the longest
survey history (Figure~\ref{fig-viability-supp} B; population 70550,
surveyed for 16 years).

@@

Disease-induced population declines are impacting global biodiversity
\citep{daszak2000}, and strategies to recover affected species are
lacking. Here, we used populations of resistant/tolerant individuals to
reestablish extirpated populations, despite ongoing disease. Although
previously proposed \citep{mendelson2019}, this strategy has rarely been
successfully implemented (see \citep{joseph2018} for a small-scale
example). In particular, although numerous studies have documented
signatures of selection in hosts following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (i) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease (Bd) and that some
of the genomic regions under selection are associated with genes related
to disease resistance/tolerance, and (ii) reintroduction of individuals
from these resistant/tolerant populations allows the successful
reestablishment of viable frog populations despite ongoing Bd infection.
Although these results are generally consistent with expectations under
evolutionary rescue \citep{carlson2014}, obtaining strong evidence in
wild animal populations for each of the associated causal links (i.e.,
selection for adaptive alleles \(\rightarrow\) increase in host fitness
in the presence of the pathogen \(\rightarrow\) positive population
growth and recovery) is difficult and perhaps impossible. Nonetheless,
in the following narrative, we use evolutionary rescue as an
informative, albeit tentative framework.

@@

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
specific alleles and genomic regions showing signatures of selection
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the MHC plays
an important role in immunity. In our study, we identified a region that
shows evidence of selection in recovering populations and contains eight
genes associated with either the MHC Class I or Class III regions. These
results corroborate numerous previous studies linking MHC genes to
amphibian resistance/tolerance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest indication of directional selection (as measured by
\(pi_{diff}\)) contains the gene GVINP1. Several previous studies have
found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018, ellison2014}. This gene was also strongly
linked to disease in salmon, explaining a remarkable 20\% of the
resistance phenotype \citep{robledo2020, robledo2018}. Finally, the RIN3
gene showed a consistent pattern of allele frequency differences across
all nine of the frog populations sampled in this study, indicating
consistent selection in populations distributed across a wide geographic
area. In the SI, we provide additional details on these and other genes
identified in our study as under selection .

@@

The recent emergence of Bd worldwide has contributed to the decline of
hundreds of amphibian species, some of which are now extinct in the wild
\citep{scheele2019}. This unprecedented impact on global amphibian
biodiversity is compounded by the lack of any effective and broadly
applicable strategies to reverse these impacts
\citep{garner2016, knapp2022}. Importantly, in addition to the natural
recovery documented for MYL frogs \citep{knapp2016}, other species are
also showing evidence of post-epizootic recovery in the presence of Bd
\citep{scheele2017, voyles2018} and suggest the possibility of using
animals from these recovering populations to reestablish extirpated
populations. As with MYL frogs, the feasibility and long-term success of
such efforts will depend on the availability of robust donor populations
containing individuals that have acquired via evolution the adaptive
alleles necessary to allow frog survival and population growth in the
presence of Bd. Despite the hopeful example of successful
reestablishment of MYL frogs via evolutionary rescue and
reintroductions, the challenge of recovering hundreds of Bd-impacted
amphibians worldwide is a daunting prospect. Although we now have a
proven strategy to reestablish extirpated MYL frog populations, recovery
across their large historical range will require substantial resources
over many decades. The results of this study provide a hopeful starting
point for that endeavor, and other future efforts worldwide.

Evolutionary rescue may be increasingly critical for the survival of
wildlife populations in a rapidly changing world. The possibility - and
importance - of evolutionary rescue has been discussed for a wide
variety of taxa \citep[e.g.,][]{mills2018, oziolor2019}, and in the
specific context of disease-impacted systems \citep{searle2020}.
However, obtaining evidence in wild animal populations for all the
causal links in this process (i.e., selection for adaptive alleles
\(\rightarrow\) increase in host fitness in the presence of the pathogen
\(\rightarrow\) positive population growth and recovery) is difficult
and perhaps impossible. Nonetheless, we provide a compelling example
from the wild of how in the face of a novel stressor evolutionary rescue
can produce individuals that harbor adaptive alleles and that these
individuals can be used to reestablish extirpated imperiled populations.
We expect that such ``assisted evolutionary rescue'' will be an
essential tool in wildlife conservation in an era of global change.

@@

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome. We then used the WINGEN R package
\citep{bishop2023} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the fitted spline, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015} . This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for F\textsubscript{ST} or above/below the mean for \(\pi_{diff}\).
Outlier regions were then identified as those that were outliers in both
analyses, meaning that they showed 1) high differentiation between naive
and recovering populations, and 2) differential patterns of nucleotide
diversity in the same region. Finally, we extracted gene transcripts
mapped within each region and retrieved annotation for that region using
blast.

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
\citep[see SI for details]{boyle2004}. To assess the magnitude of
changes in Bd load on frogs following translocation, we compared Bd
loads measured before versus after translocation. Before-translocation
loads were quantified using skin swabs collected from all
to-be-translocated frogs at the donor site on the day before or the day
of the translocation. After-translocation Bd loads were based on all
swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above. Briefly,
the response variable was before-translocation load and the predictor
variable was after-translocation load. We compared models with and
without group-level effects, and group-level effects included
pit\_tag\_id, pit\_tag\_id nested within site\_id, or pit\_tag\_id
nested within translocation\_id. All models used a negative binomial
family. Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the
following GitHub repository:
\url{https://github.com/SNARL1/translocation}.

@@

After parameterizing our model with CMR-estimated adult frog survival
probabilities and other known vital rates (Table S2), we performed four analyses (described in detail in the SI).
First, we compute the long-run growth rate \(\lambda\) for each of our
12 translocated populations to determine if the populations were
deterministically predicted to growth or decline in the long-run.
Second, we compute the elasticity of \(\lambda\) to model parameters, to
quantify how much changes in adult survival, for example, affected the
long-run growth rate. Third, we included demographic and environmental
stochasticity into our model and simulated the 50-year viability (i.e.,
1 - extinction probability) of populations given an introduction of 40
adult individuals into an unoccupied habitat. Finally, we fit our model
to our longest translocation trajectory to confirm that our model could
reasonably reproduce the observed recovery trajectories of MYL frogs
following reintroductions.}
